---
title: "Exercise1_Eastaury"
author: "sumuthuni"
date: "2026-08-19"
output:
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE
)
```

# Introduction

This exercise develops a reproducible workflow for cleaning, integrating, analysing, and visualising estuarine ecological datasets. Fish catch records, water-quality data, spatial metadata, and species taxonomy were combined to examine fish abundance along a salinity gradient.

# Load Packages

```{r}
library(tidyverse)
library(readxl)
library(lubridate)
library(stringr)
library(here)
library(knitr)
library(kable)


```

# Run Analysis

```{r}
source(here::here("code", "Exercise_estauries.R"))
```

# Clean Master Dataset

```{r}
head(master_dataset, 10) |>
  Knit::kable(
    digits = 2,
    caption = "First 10 rows of the cleaned estuary master dataset"
  )
```

# Statistical Summary Table

```{r}
summary_table |>
  kable(
    digits = 2,
    caption = "Fish abundance and salinity summary by species and estuary zone"
  )
```

# Fish Abundance Along the Salinity Gradient

```{r fig.width=10, fig.height=7}
estuary_plot
```

# Conclusion

The workflow successfully cleaned and integrated the estuary datasets into a single master dataset. The resulting summary table and multi-panel plot demonstrate variation in fish abundance across salinity gradients and estuary zones.
