---
title: "Workshop2"
author: "sumuthuni"
date: "2026-07-19"
output: html_document
---
Practical Exercises: Penguins, Pivots, and Relational Data
# Make sure packages are loaded
library(tidyverse)
library(palmerpenguins)

# The messy metadata provided by your colleague
island_metadata <- tibble(
  island_name = c(" biscoe", "Dream ", "Torgersen"),
  station_install = c("15/01/2003", "22-03-2004", "05/11/2001"),
  latitude = c(-64.81, -64.73, -64.76)
)

print(island_metadata)

#graph
library(ggplot2)

## Scroll to the right to see the new columns
head(penguins_spatial)

# Load packages
library(dplyr)
library(stringr)
library(lubridate)

## Exercise 1: Clean the metadata
clean_metadata <- island_metadata |>
  mutate(
    island_name = str_trim(island_name),
    island_name = str_to_title(island_name),
    station_install = dmy(station_install)
  )
  ## Check cleaned metadata
print(clean_metadata)

# Exercise 2: Join metadata with penguins dataset
penguins_spatial <- penguins |>
  left_join(
    clean_metadata,
    by = join_by(island == island_name)
  )

# Check the result
head(penguins_spatial)

#Graph box plot

  #Boxplot
  ggplot(penguins_spatial,
       aes(x = island,
           y = body_mass_g,
           fill = island)) +
  geom_boxplot() +
  labs(
    title = "Body Mass Distribution by Island",
    x = "Island",
    y = "Body Mass (g)"
  ) +
  theme_minimal()
#Exercise 3
# Load required packages
library(dplyr)
library(tidyr)

# Create a wide summary table
body_mass_summary <- penguins_spatial |>
  drop_na(body_mass_g) |>
  group_by(species, island) |>
  summarise(
    max_body_mass = max(body_mass_g),
    .groups = "drop"
  ) |>
  pivot_wider(
    names_from = island,
    values_from = max_body_mass
  )

# View the table
print(body_mass_summary)
```{r}
#creating table
library(knitr)
body_mass_summary |>
  kable(
    caption = "Maximum Body Mass (g) of Penguin Species by Island"
  )
```{r}
library(dplyr)
library(tidyr)
library(knitr)

body_mass_summary <- penguins_spatial |>
  drop_na(body_mass_g) |>
  group_by(species, island) |>
  summarise(
    max_body_mass = max(body_mass_g),
    .groups = "drop"
  ) |>
  pivot_wider(
    names_from = island,
    values_from = max_body_mass
  )

kable(
  body_mass_summary,
  caption = "Maximum Body Mass (g) of Penguin Species by Island"
)
```
